---
title: "kms: foRmulas foR keRas"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{kerasformula}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
type: docs
repo: https://github.com/rstudio/keras
menu:
  main:
    name: "kms: foRmulas foR keRas"
    identifier: "keras-R-formulas"
    parent: "keras-using-keras"
    weight: 50
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(comment = "", message = FALSE, warning = FALSE)
```


The goal of this document is to introduce `kms` (as in `keras_model_sequential()`), a regression-style function which allows users to call `keras` neural nets with `R` `formula` objects (hence, library(`kerasformula`)). 

First, make sure that `keras` is properly configured:

```{r, eval = FALSE}
install.packages("keras")
library(keras)
install_keras() # see https://keras.rstudio.com/ for details. 
library(kerasformula)
```

`kms` splits training and test data into sparse matrices.`kms` also auto-detects whether the dependent variable is categorical or binary. `kms` accepts the major parameters found in `library(keras)` as inputs (loss function, batch size, number of epochs, etc.) and allows users to customize basic neural nets (dense neural nets of various input shapes and dropout rates). The final example below also shows how to pass a compiled `keras_model_sequential` to `kms` (preferable for more complex models). 

This example works with some of the `imdb` movie review data that comes with library(`keras`). Specifically, this example compares the default dense model that `ksm` generates to the `lstm` model described [here](https://keras.rstudio.com/articles/examples/imdb_lstm.html). To expedite package building and installation, the code below is not actually run but can be run in under six minutes on a 2017 MacBook Pro with 16 GB of RAM (of which the majority of the time is for the lstm). 

```{r, eval = FALSE}
max_features <- 5000 # 5,000 words (ranked by popularity) found in movie reviews
maxlen <- 50  # Cut texts after 50 words (among top max_features most common words) 
Nsample <- 1000 

cat('Loading data...\n')
imdb <- keras::dataset_imdb(num_words = max_features)
imdb_df <- as.data.frame(cbind(c(imdb$train$y, imdb$test$y),
                               pad_sequences(c(imdb$train$x, imdb$test$x))))

set.seed(2017)   # can also set kms(..., seed = 2017)

demo_sample <- sample(nrow(imdb_df), Nsample)
P <- ncol(imdb_df) - 1
colnames(imdb_df) <- c("y", paste0("x", 1:P))

out_dense <- kms("y ~ .", data = imdb_df[demo_sample, ], Nepochs = 10)

plot(out_dense$history)  # incredibly useful 
# choose Nepochs to maximize out of sample accuracy

out_dense$confusion
```


```
    1
  0 107
  1 105
```
```{r, eval=FALSE}
cat('Test accuracy:', out_dense$evaluations$acc, "\n")
```
```
Test accuracy: 0.495283 
```

Pretty bad--that's a 'broken clock' model. Suppose want to add some more layers. 

```{r, eval = FALSE}
out_dense <- kms("y ~ .", data = imdb_df[demo_sample, ], Nepochs = 10, 
                 layers = list(units = c(512, 256, 128, NA), 
                               activation = c("relu", "relu", "relu", "softmax"), 
                               dropout = c(0.5, 0.4, 0.3, NA)))
out_dense$confusion
```
```
     1
  0 113
  1 105
```
```{r, eval = FALSE}
cat('Test accuracy:', out_dense$evaluations$acc, "\n")
```
```
Test accuracy: 0.4816514
```

No progress. Suppose we want to build an `lstm` model and pass it to `ksm`.

```{r, eval = FALSE}
k <- keras_model_sequential()
k %>%
  layer_embedding(input_dim = max_features, output_dim = 128) %>% 
  layer_lstm(units = 64, dropout = 0.2, recurrent_dropout = 0.2) %>% 
  layer_dense(units = 1, activation = 'sigmoid')

k %>% compile(
  loss = 'binary_crossentropy',
  optimizer = 'adam',
  metrics = c('accuracy')
)
out_lstm <- kms(input_formula = "y ~ .", data = imdb_df[demo_sample, ], 
                keras_model_seq = k, Nepochs = 5, seed = 12345)
out_lstm$confusion
```
```
    0  1
  0 80 17
  1 25 77
```

```{r, eval=FALSE}
cat('Test accuracy:', out_lstm$evaluations$acc, "\n")
```
```
Test accuracy: 0.7889447 
```

78.9% out-of-sample accuracy. That's marked improvement!

For another worked example starting with raw data (from `rtweet`) visit [here](https://github.com/rdrr1990/code/blob/master/kms.md).