---
title: "compatability check for kerasformula"
author: "Pete Mohanty"
date: "5/2/2019"
output: github_document
---

This document attempts to fit a simple neural net using `kerasformula`. It also displays a number of version checks on `kerasformula` and its dependencies both in `R` and in `Python`. This document is intended to knit whether or not the model can be fit successfully. To run this code on your machine, [click here](https://raw.githubusercontent.com/rdrr1990/kerasformula/master/short_course/kerasformula_diagnostic.Rmd).

```{r setup}
if(!require(keras)) install.packages(keras)
if(!require(kerasformula)) install.packages(kerasformula)
library(keras)
if(!is_keras_available()) install_keras()
keras_installed <- is_keras_available()
```


```{r example, fig.height=3}
if(keras_installed){
  library(kerasformula)
  tried <- try(out <- kms(Species ~ ., iris, verbose=0))
  if(!inherits(tried, "try-error")) plot_confusion(out)
}else{
  message("keras did not install properly.\n")
}

```

```{r}
system("which python > python_path.txt")
python_path <- readLines("python_path.txt")
python_path
```

If that path is correct, it should be set in two different ways. Making `changing_settings=TRUE` would accomplish that. 

```{r}
if(!require(reticulate)) install.packages(reticulate)
library(reticulate)

change_settings <- FALSE

if(change_settings){
  
  System.setenv(TENSORFLOW_PYTHON=python_path)
  use_python(python_path)
    
}
```
The Python path should appear for each of these key libraries...
```{r}
py_discover_config("tensorflow")
py_discover_config("numpy")
py_discover_config("keras")
```

```{r}
sessionInfo()

```
